Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-06-09 mu/article/article.tex
Introduction completed.

Since its discovery, graphene has attracted a great interest in materials
science and condensed matter physics being a new class of materials that are
only one atom thick\cite{geimNM07}. This allotrope of carbon with hexagonal
lattice structure is almost transparent and presents properties like fractional
quantum Hall effect at room temperature, excellent thermal transport
properties, high conductivity and strength \cite{geimNM07, reinaNL08,
novoselov2S07, balandinNL08} and is a perfect platform to be used in two-
dimensions electronic systems; however most electronic applications are
disabled by the absence of a semiconducting gap. Recent studies demonstrate
that the band gap of graphene can be opened by applying an electric
field\cite{zhangN09}, reducing the surface area\cite{hanPRL07} or applying
uniaxial strain\cite{niACSN08}. Another option to open the gap is by doping;
this has been successfully achieved using boron and nitrogen\cite{guoIJ11} and
hydrogen\cite{eliasS09, guisingerNL09, samarakoonACSN10}.

Hydrogenated graphene can result in different structures depending of the
percentage of hydrogenation and the position at which they are bonded in the
hexagonal lattice. In this paper we present a study made on two 50\%
hydrogenated graphene structures, \emph{up} and \emph{alt} , both showing a
opening of the gap. As shown in Figs. \ref{fig:up-struc} and 
\ref{fig:alt-struc} the \emph{up} structure has hydrogen atoms bonded to the
carbon layer only in the upper side of the structure while the \emph{alt}
structure has hydrogen alternating in the upper and bottom sides of the carbon
slab.

We conduct a theoretical study of the spin velocity injection, being this a
nonlinear response, in both structures. We analyze particularly the cases when
the spin is polarized along a Cartesian direction or when the velocity is
directed along the $x$ or $y$ directions on the $xy$ plane of the structures.

@@

>>>>

{
\color{blue}
Spintronics is an emerging research field of electronics in which the
manipulation and transport of spin of electrons in a solid state 
media plays the determining role adding a new degree of freedom to the
conventional charge manipulation.\cite{wolfSC04,fabianAPS07}
% 
At present there is an increasing interest in attain the same level of control
over the transport of spin at micro or nano scales as has been done for the flow
of charge in typical electronic devices.\cite{awschalomNP2007} Some
semiconductor spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90} and some of them require spin polarized electrical current
\cite{awschalomSSBM13} or pure spin currents (PSC). In PSCs there is no net
motion of charge; spin-up electrons move in a given direction while spin-down
electrons travel in the opposite one. This phenomena can result from spin
injection,\cite{malPRB03} Hall Effects,\cite{sinovaPRB04} interference of two
optical beams,\cite{bhatPRL00, najmaiePRB03} or one photon absorption of
linearly polarized light\cite{bhatPRL05} and has been observed in
gallium arsenide (GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium
arsenide (AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

% Since its discovery, graphene has attracted a great interest in materials
% science and condensed matter physics being a new class of materials that are
% only one atom thick\cite{geimNM07}. 
% 
Graphene, an allotrope of carbon with hexagonal 2D lattice structure presents
properties like fractional quantum Hall effect at room temperature, excellent
thermal transport properties, high conductivity and strength \cite{geimNM07,
reinaNL08, novoselov2S07, balandinNL08} being then a perfect platform to be used
in two-dimensions electronic systems; however most electronic applications are
disabled by the absence of a semiconducting gap. Recent studies demonstrate that
the band gap of graphene can be opened by applying an electric
field,\cite{zhangN09} reducing the surface area,\cite{hanPRL07} or applying
uniaxial strain.\cite{niACSN08} Another possibility to open the gap is by
doping; this has been successfully achieved using nitrogen,\cite{weiNL2009}
boron-nitrogen,\cite{guoIJ11} silicon,\cite{colettiPRB10} noble-metals, and
hydrogen.\cite{eliasS09, guisingerNL09, samarakoonACSN10}
% 
Depending on the percentage of hydrogenation and spatial configurations of
hydrogen-carbon bonds, hydrogenated graphene can result in different spatial
configurations.
% 
In this paper we present two 50\% hydrogenated graphene non centrosymmetric
structures both presenting a discernible band gap: the \emph{up} structure,
shown in Fig. \ref{fig:up-struc}, has hydrogen atoms bonded to the carbon layer
only in the upper side of the structure while the \emph{alt} structure, shown in
Fig. \ref{fig:alt-struc}, has hydrogen alternating in the upper and bottom sides
of the carbon slab.\cite{zapataPSB2016}

Using those structures we address a theoretical study of the spin velocity
injection (SVI) by one-photon absorption of linearly polarized light making
emphasis for the particular cases when the spin of electrons is directed along
the the $z$ Cartesian coordinate, perpendicular to the $xy$ plane of the
structure, or for the case when the velocity is directed along the $x$ and $y$
Cartesian direction on the $xy$ plane of the structures.
% 
The SVI ($\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$) is a nonlinear optical
effect that quantifies the velocity at which a PSC moves along the Cartesian
direction $\mathrm{b}$ with the spin of electron polarized along the Cartesian
direction $\mathrm{a}$. One photon absorption of linearly polarized light can
promote a distribution of electrons in $\mathbf{k}$ space regardless the
symmetry of the material resulting in a not net electrical
current. Then, the electrons excited to the conduction bands at
opposite $\mathbf{k}$ points will result in opposite spin polarizations 
producing no net spin injection.\cite{bhatPRL05} If the crystalline structure of
the material is not centrosymmetric the spin polarization injected at a given k
point could not vanish\cite{alvaradoPRL85, schmiedeskampPRL88} and then a PSC
will be produced since the velocities of electrons at opposite $\mathbf{k}$
points are opposite.

This paper is organized as follows. In Section \ref{sec:theory} we present the
theory and formulas that describe PSC and SVI. In Section \ref{sec:results} we
describe the details of calculations and the corresponding SVI spectra for the 
\emph{up} and \emph{alt} structures. Finally, in Section

}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%                          %%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%   C O N C L U S I O N S  %%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%                          %%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions} % (fold)
\label{sec:conclusions}
